---
title: "Multiple linear regression"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r load-data, eval=TRUE}
load("more/evals.RData")
```


## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.  
    
<b>This is an observational study. The question should be rephrased as 'is the beauty of a professor a predictor of score in course evaluations?</b>  


2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?  
    
```{r message=FALSE,warning=FALSE}

library(tidyverse)
ggplot(evals,aes(x=score))+
  geom_histogram(aes(y=..count..,fill=..count..,alpha=.2))+
  scale_fill_gradient("Count", low="blue", high="red")+
  theme_classic()+
  theme(legend.position="none")

```
    
<b>The distribution is left-skewed. Positive evaluations are more frequent than negative evaluations; a normal distribution was expected, wherein fewer positive evaluations were submitted.</b>  


3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).  
    
```{r message=FALSE,warning=FALSE}

evals$agecat[evals$age<=40] <-1
evals$agecat[evals$age>=40 & evals$age<=50] <-2
evals$agecat[evals$age>=50 & evals$age<=60] <-3
evals$agecat[evals$age>=60] <-4
evals$agecat<-as.numeric(evals$agecat)

ggplot(evals,aes(y=evals$bty_avg,x=as.factor(evals$agecat)))+
  geom_boxplot(color="black",fill="#80bfff",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5) +
  scale_x_discrete(labels=c("1"="40 or younger",
                              "2"="41 - 50",
                              "3"="51 - 60",
                              "4"="61 or older"))

```

<b>There is a decline in median beauty as age increases when evaluations are categorized by age range.</b>  


## Simple linear regression  

```{r scatter-score-bty_avg, eval = FALSE}
plot(evals$score ~ evals$bty_avg)
```


Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?  

<b>The plotted points are superimposed; it appears that there are fewer plotted points than there should be.</b>  


4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?  
    
```{r message=FALSE,warning=FALSE}

ggplot(evals, aes(bty_avg, score))+
  geom_point(position = position_jitter(w = 0.1, h = 0.1))+
  theme_classic()+
  ylab("score")+
  xlab("beauty")

```

5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?  
    
```{r message=FALSE,warning=FALSE}

m_bty <- lm(evals$score ~ evals$bty_avg)
ggplot(evals, aes(bty_avg, score))+
  geom_point(position = position_jitter(w = 0.1, h = 0.1))+
  geom_abline(intercept = coef(m_bty)["(Intercept)"],
              slope = coef(m_bty)["evals$bty_avg"])+
  theme_classic()+
  ylab("score")+
  xlab("beauty")


```

The equation is $\hat{y}$=`r coef(m_bty)["(Intercept)"]` + `r coef(m_bty)["evals$bty_avg"]`(bty_avg)  

```{r}
summary(m_bty)
```
<b>`bty_avg` is indeed a predictor of evaluation score (p < 0.001), however the slope of the equation indicates that for every point increase in beauty there is a fractional increase in evaluation score (.06664).  


6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).  
    
```{r}

ggplot(evals, aes(y=m_bty$residuals, x=evals$bty_avg))+
  geom_point()+
  geom_abline(intercept = 0,
              slope = 0)+
  theme_classic()

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(x=m_bty$residuals))+
  geom_histogram(aes(y=..count..,fill=..count..,alpha=.2))+
  theme_classic()+
  theme(legend.position="none")

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(sample=m_bty$residuals))+
  stat_qq()+
  stat_qq_line()+
  theme_classic()

```

<b>While the residuals appear to confirm constant variability, the histogram and the normal probability plot both suggest a distribution that deviates from a normal distribution.</b>  


## Multiple linear regression  


```{r bty-rel}

m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)

```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.  
    
```{r message=FALSE,warning=FALSE}

ggplot(evals, aes(y=m_bty_gen$residuals, x=m_bty_gen$fitted.values))+
  geom_point()+
  geom_abline(intercept = 0,
              slope = 0)+
  theme_classic()

```

<b>Plot of residuals against fitted values indicates constant variance.</b>  


```{r message=FALSE,warning=FALSE}

ggplot(evals, aes(y=m_bty_gen$residuals, x=evals$bty_avg))+
  geom_point()+
  geom_abline(intercept = 0,
              slope = 0)+
  theme_classic()

```

<b>Variance appears constant.</b>  


```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(sample=m_bty_gen$residuals))+
  stat_qq()+
  stat_qq_line()+
  theme_classic()

```

<b>The distribution is approximiately normal. Conditions for the regression are reasonable.</b>  


8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?  
    
<b>The inclusion of the variable improved the model; `bty_avg` is still a sgnificant predictor of `score`.</b>  


```{r twoLines, eval = FALSE}
multiLines(m_bty_gen)
```

9.  What is the equation of the line corresponding to males? (*Hint:* For
    males, the parameter estimate is multiplied by 1.) For two professors
    who received the same beauty rating, which gender tends to have the
    higher course evaluation score?  
    
\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (1) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \end{aligned} 
\]  

<b>The additional term in the equation implies that males would have the higher course evaluation if beauty rating is held constant.</b>  


10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.  
    
```{r message=FALSE,warning=FALSE}

m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
multiLines(m_bty_rank)

```

<b>R creates two variables, `ranktenure track` and `ranktenured`.</b>  


## The search for the best model  


11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.

<b>I would expect `cls_profs` to have the least association with evaluation score.</b>  


```{r m_full, eval = FALSE, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.  
    
<b>`cls_profs` has the greatest p-value, with `cls_profssingle` p-value at 0.77806.</b>  

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$cls_profs)))+
  geom_boxplot(color="black",fill="#80bfff",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

13. Interpret the coefficient associated with the ethnicity variable.  

<b>The coefficient indicates that professors not identified as a minority have higher evaluation scores, but the p-value indicates negligible significance.</b>  


14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?  
    
```{r message=FALSE, warning=FALSE}

m_full2 <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full2)

```

<b>The coefficients and significance changed; each variable is more significantly associated.</b>  


15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.  
    
```{r message = FALSE,warning=FALSE}

m_full_final <- lm(score ~ ethnicity + gender + language + age + cls_perc_eval 
              + cls_credits + bty_avg + pic_color, data = evals)
summary(m_full_final)

```

Equation: $\hat{score}$ = `r coef(m_full_final)["(Intercept)"]` + 
`r coef(m_full_final)["ethnicitynot minority"]`(ethnicitynot minority) + 
`r coef(m_full_final)["gendermale"]`(gendermale) + 
`r coef(m_full_final)["languagenon-english"]`(languagenon-english) + 
`r coef(m_full_final)["age"]`(age) + 
`r coef(m_full_final)["cls_perc_eval"]`(cls_perc_eval) + 
`r coef(m_full_final)["cls_creditsone credit"]`(cls_creditsone credit) + 
`r coef(m_full_final)["bty_avg"]`(bty_avg) + 
`r coef(m_full_final)["pic_colorcolor"]`(pic_colorcolor)

16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.  
    
```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(sample=m_full_final$residuals))+
  stat_qq()+
  stat_qq_line()+
  theme_classic()

```


```{r message=FALSE,warning=FALSE}

ggplot(evals, aes(y=m_full_final$residuals, x=m_full_final$fitted.values))+
  geom_point()+
  theme_classic()

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$ethnicity)))+
  geom_boxplot(color="black",fill="#80bfff",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$gender)))+
  geom_boxplot(color="black",fill="#80baaf",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$language)))+
  geom_boxplot(color="black",fill="#60eaaf",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=evals$age))+
  geom_point(shape=19,color="blue",fill="#efccff")+
  theme_classic()+
  theme(legend.position="none")+
  scale_x_continuous(name="cls_perc_eval",
                   breaks=c(0,20,40,60,80,100))

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=evals$cls_perc_eval))+
  geom_point(shape=19,color="red",fill="#efccff")+
  theme_classic()+
  theme(legend.position="none")

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$cls_credits)))+
  geom_boxplot(color="black",fill="#60eaaf",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=evals$bty_avg))+
  geom_point(shape=19,color="gray",fill="#efccff")+
  theme_classic()+
  theme(legend.position="none")

```

```{r message=FALSE,warning=FALSE}

ggplot(evals,aes(y=evals$score,x=as.factor(evals$pic_color)))+
  geom_boxplot(color="black",fill="#60ffff",alpha=0.2)+
  theme_classic()+
  theme(legend.position="none")+
  geom_text(aes(label = ..count.., y= ..prop..), stat= "count", vjust = -.5)


```

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?  
    
<b>This information should not change any of the conditions of linear regression.</b>  


18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.  
    
<b>The highest scores are associated with younger, attractive (high rank in beauty score) males who do not identify as a minority and teach a course for one credit.</b>  


19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?  
    
<b>These conclusions would not apply to all universities and professors. University demographics would likely greatly interfere with our model.</b>  


